---
title: "Exploration of milestone inscriptions in the EDCS dataset"
author: "Petra Hermankova"
date: "01/10/2020"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged 
 
---

This notebook explores all the milestones in the **Epigraphic Database Clauss-Slaby** (http://www.manfredclauss.de/, henceforth EDCS). The primary aim is to:

- find all inscriptions categorized as 'milestones'
- provide basic descriptive statistics: how many there are, what categories of monuments we have, what materials they are made from, where do they come from etc. 
- display all the milestones on the map along with the map of Roman provinces and Roman roads

The script takes the reader through all the steps of selection, descriptive statistics and through the various displays on the interactive map. Address all the questions and comments to `petra.hermankova@cas.au.dk`.

# Initial setup

## Setup of the environment:

```{r setup, echo=TRUE, message=FALSE}
devtools::install_github("sdam-au/sdam") # loading SDAM custom package, if not working try devtools::install_github("mplex/cEDCSar", subdir="pkg/sdam")
#devtools::install_github("mplex/cEDCSar", subdir="pkg/sdam")
library(tidyverse)
library(sdam)
library(jsonlite)
library(getPass)
library(formatR)
library(leaflet)
```

## Loading data
Load the dataset, if you have Sciencedata.dk credentials

```{r, echo=FALSE}
mycred_secret<- readLines("~/mysecret.txt")
```

```{r, loading data}
resp2 = request("EDCS_text_cleaned_2021-03-01.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDCS/public", method="GET", cred=mycred_secret)
```

```{r, echo=FALSE}
remove(mycred_secret)
```

OR of you don't have Sciencedata credentials:
```{r, loading data public, eval=FALSE}
resp2 = request("EDCS_text_cleaned_2021-03-01.json", path="/public/1f5f56d09903fe259c0906add8b3a55e/", method="GET", anonymous = TRUE, cred = NULL)
```

Make a list and tibble from the downloaded dataset
```{r}
list_json2 <- jsonlite::fromJSON(resp2)
EDCS_tibble = as_tibble(list_json2)
```

## Exploring the data
Display the first six records of the dataset
```{r}
head(EDCS_tibble)
```

# Finding all the milestones 
What attributes (columns) might contain milestone related information? Let's display all the the attribute names.
```{r}
EDCS_tibble %>% 
  names()
```

Based on the domain expertise and previous explorations of the dataset, any milestone related information is most likely contained in the following attributes: `inscr_type` and `comment`.

## Milestone as type of inscription

```{r}
unique(unlist(EDCS_tibble$inscr_type))
```

```{r}
EDCS_insc_type<- unlist(EDCS_tibble$inscr_type)

EDCS_insc_type %>% 
  table() %>% as.data.frame -> EDCS_insc_type
EDCS_insc_type
```

### How many inscriptions are categorised as a milestone?
```{r}
milestone_EDCS <- EDCS_tibble %>% 
  filter(inscr_type %in% str_subset(inscr_type, "miliaria"))
nrow(milestone_EDCS)
```

## Milestone in the `Comment` attribute
Free text commentary also contains infromation about milestones, however it is not structired in any way. 
Inspect the commentary for milestone related comments. Note: commentary is mostly written in German or English.

### How many inscriptions contain any milestone-related commentary?
```{r}
str_subset(EDCS_tibble$Comment, "[m|M]ilest*")
str_subset(EDCS_tibble$Comment, "[m|M]eilen*")
```

# Descriptive overview 

## How many milestones total are in the EDCS dataset when we combine all the search criteria?
```{r}
nrow(milestone_EDCS)
```

## How many % of the total number of EDCS inscriptions represent milestones
```{r}
nrow(milestone_EDCS)/(nrow(EDCS_tibble)/100)
```

## Type of inscription

What different types of inscriptions belong to the discussed group of milestones?

```{r}
unique(milestone_EDCS$inscr_type)
```

## Language

What language were written the inscriptions belonging to the discussed group of milestones? NA == Latin
```{r}
unnest_auto(milestone_EDCS, language) %>% 
  count(language, sort=TRUE)
```


## Material 
What different types of material is represented in the discussed group of milestones?

```{r}
milestone_EDCS %>% 
  count(Material, sort=T)
```


## Milestones present in EDH
Some milestones might be present in EDCS and in EDH as well.

```{r}
milestones_EDCS_EDH<- milestone_EDCS %>% 
  filter(Links %in% str_subset(EDCS_tibble$Links, "inschrift\\/HD\\d{5,6}")) %>% 
  mutate(EDH_id = str_extract(milestones_EDCS_EDH$Links, "HD\\d{5,6}"))

library(tidyr)
unnest_longer(milestones_EDCS_EDH, inscr_type) %>% 
  count(inscr_type, sort= T)

```

# Text exploration

```{r}
library(tidytext)

milestone_EDCS_text<- milestone_EDCS %>% 
  unnest_tokens(word, clean_text_interpretive_word)
```

## How many words total
```{r}
nrow(milestone_EDCS_text)
```

## The most common words

```{r}
milestone_EDCS_text %>% 
  count(word, sort=T)
```


## N-grams

### Bigram
```{r}
milestone_EDCS_bigram<- milestone_EDCS %>% 
  unnest_tokens(output = bigram, input = clean_text_interpretive_word, drop = FALSE, token = "ngrams", n = 2, collapse = FALSE)
```

```{r}
milestone_EDCS_bigram %>% 
  count(bigram, sort=T)
```

### Trigram
```{r}
milestone_EDCS_trigram<- milestone_EDCS %>% 
  unnest_tokens(output = bigram, input = clean_text_interpretive_word, drop = FALSE, token = "ngrams", n = 3, collapse = FALSE, to_lower = FALSE)

milestone_EDCS_trigram %>% 
  count(bigram, sort=T)
```

## Detect all Latin numerals

```{r}
library(googlesheets4)

gs4_deauth() # de-uthorized mode, no need of authentication token (if the spreadsheet is public)
numerals<- read_sheet("https://docs.google.com/spreadsheets/d/1RKRNMlSjB3yF3FHXPLhGfnLJis63300R9x65BdAKa8o/edit?usp=sharing", sheet = "Numerals")

numerals
```

Latin numerals on milestones:
```{r}
milestone_EDCS_text %>% 
  filter(word %in% numerals$Latin_num) %>% 
  count(word, sort = T)
```

Greek numerals on milestones:
```{r}
milestone_EDCS_text %>% 
  filter(word %in% numerals$Greek_num) %>% 
  count(word,  sort = T)
```

### Detecting distance in Roman miles

```{r}
distances<- as.data.frame(str_extract(milestone_EDCS$clean_text_interpretive_word, "(?<=milia passuum ).+"))
distances <- distances %>% 
  rename(extracted = `str_extract(milestone_EDCS$clean_text_interpretive_word, "(?<=milia passuum ).+")`)
unique(distances$extracted)
```






## Building a stopword list
```{r}
stopwords<- c("et", "a", "ab", "ad", "per", "ac", "in", "καὶ", "τὸ", "ἀπὸ", "ἡ")
```






















# Spatial exploration

## Ancient findspot

What kind of information do we have about ancient findspot of a milestone? Let's see first 20 records.
```{r}
milestone_EDCS$place[1:20]
```

How many different ancient findspots do we have, including "NULL"?
```{r}
length(unique(milestone_EDCS$place))
```

## Distribution by Roman provinces
How many milestones were found in individual Roman provinces? Display from the province containing the most to least milestones.
```{r}
milestone_EDCS %>% 
  count(province, sort =TRUE)
```

Display all the provinces and milestones in a chart: 
```{r, fig.height=10, fig.width=6}
milestone_EDCS %>% 
  count(province) %>% 
  ggplot(aes(y=fct_rev(province), x=n)) +
  geom_point(color="red") +
  #coord_fixed(ratio = 7/1) +
  labs(x = "Number of milestones", y = "Roman province", title = "Number of milestones per Roman Province", subtitle = ggtitle(paste("n =", nrow(milestone_EDCS)), "inscriptions")) +
    theme_linedraw(base_size = 12)
```

## Map of all milestones (with available coordinates)
```{r}
milestone_EDCS
```

```{r}

# milestones with no coordinates
milestone_EDCS_nocoord<- milestone_EDCS %>% 
  dplyr::filter(is.na(Latitude) | is.na(Longitude))

# milestones with coordinates
milestone_EDCS_coord<- milestone_EDCS %>% 
  dplyr::filter(!is.na(Latitude) | !is.na(Longitude))

milestone_EDCS_coord$Latitude <- as.numeric(milestone_EDCS_coord$Latitude)
milestone_EDCS_coord$Longitude <- as.numeric(milestone_EDCS_coord$Longitude)
```

### How many milestones have spatial coordinates
```{r}
nrow(milestone_EDCS_coord)

nrow(milestone_EDCS_coord)/(nrow(milestone_EDCS)/100)
```
                   
## Milestones as dots
```{r}
map_milestones_EDCS <-leaflet(width="100%") %>%
  addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
  addCircles(lng = milestone_EDCS_coord$Longitude, 
             lat = milestone_EDCS_coord$Latitude, radius = 10, fill = TRUE, color= , fillColor = milestone_EDCS_coord$Material,
             popup = paste0("<b> InscriptionID: </b>", milestone_EDCS_coord$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", milestone_EDCS_coord$place,
                           "<br><b> Type of inscription: </b>", milestone_EDCS_coord$inscr_type,
                              "<br><b> Material: </b>", milestone_EDCS_coord$Material,
                              "<br><b> Not before (date): </b>", milestone_EDCS_coord$start_yr,
                             "<br><b> Not after (date): </b>", milestone_EDCS_coord$end_yr_1,
                                "<br><b> Commentary: </b>", milestone_EDCS_coord$Comment),
) %>% 
addLegend(position = "topright",
  colors = c("Blue"),
  labels = c("Milestones"), opacity = 1,
  title = "Position of all milestones (EDCS)" 
)
map_milestones_EDCS

```


## Clustering of milestones
```{r}
map_milestones_EDCS_cluster <-leaflet(width="100%") %>%
  #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
  addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  addMarkers(lng = milestone_EDCS_coord$Longitude, 
             lat = milestone_EDCS_coord$Latitude,
             popup = paste0("<b> InscriptionID: </b>", milestone_EDCS_coord$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", milestone_EDCS_coord$place,
                           "<br><b> Type of inscription: </b>", milestone_EDCS_coord$inscr_type,
                              "<br><b> Material: </b>", milestone_EDCS_coord$Material,
                              "<br><b> Not before (date): </b>", milestone_EDCS_coord$start_yr,
                             "<br><b> Not after (date): </b>", milestone_EDCS_coord$end_yr_1,
                                "<br><b> Commentary: </b>", milestone_EDCS_coord$Comment),
             
    clusterOptions = markerClusterOptions()
  ) %>% 
addLegend(position = "topright",
  colors = c("Blue", "Green", "Yellow", "Orange"),
  labels = c("Individual inscription", "Small cluster", "Medium cluster", "Large cluster"), opacity = 1,
  title = "Clustering of all milestones in the EDCS dataset" 
)
map_milestones_EDCS_cluster
```

## Milestones and Roman provinces (200 AD)

```{r, echo=FALSE}
library(rgdal)
library(raster)
library(sf)
library(sp)
```
Source of Roads shapefile: http://awmc.unc.edu/awmc/map_data/shapefiles/ba_roads/
Source of Roman provinces (AD 200): http://awmc.unc.edu/awmc/map_data/shapefiles/cultural_data/political_shading/roman_empire_ad_200/shape/

```{r, echo=FALSE}
setwd("~/Github/epigraphic_roads")
dir.create("../data/ba_roads")
unzip("../data/ba_roads.zip", exdir = "../data/ba_roads/")
dir.create("../data/roman_empire_200_ad_provinces")
unzip("../data/roman_empire_ad_200_provinces.zip", exdir = "../data/roman_empire_200_ad_provinces")
```

```{r, echo=FALSE}
roads <- st_read("../data/ba_roads/ba_roads.shp")
provinces200 <- st_read("../data/roman_empire_200_ad_provinces/roman_empire_ad_200_provinces.shp")
```

```{r, echo=FALSE}
orbis <- readLines("https://raw.githubusercontent.com/emeeks/orbis_v2/master/base_routes.geojson") %>% paste(collapse = "\n") %>% 
  fromJSON(simplifyVector = FALSE)
```

```{r}

map_milestones_EDCS_provinces<- leaflet(width="100%") %>%
 #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
 addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
 #addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
 addPolylines(data = provinces200, color = "red", weight = 2, opacity = 0.7) %>% 
  addCircles(lng = milestone_EDCS_coord$Longitude, 
             lat = milestone_EDCS_coord$Latitude, radius = 10, fill = TRUE, color = "blue" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", milestone_EDCS_coord$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", milestone_EDCS_coord$place,
                           "<br><b> Type of inscription: </b>", milestone_EDCS_coord$inscr_type,
                              "<br><b> Material: </b>", milestone_EDCS_coord$Material,
                              "<br><b> Not before (date): </b>", milestone_EDCS_coord$start_yr,
                             "<br><b> Not after (date): </b>", milestone_EDCS_coord$end_yr_1,
                                "<br><b> Commentary: </b>", milestone_EDCS_coord$Comment),
) %>% 
addLegend(position = "topright",
  colors = c("Blue", "Red"),
  labels = c("Milestones", "Roman Provinces (200 AD)"), opacity = 1,
  title = "Milestones in the Roman Empire" 
)
map_milestones_EDCS_provinces
```

## Milestones and Roman roads
```{r}
map_milestones_EDCS_roads<- leaflet(width="100%") %>%
 addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
 #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
 #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  setView( lng = 22.326743, lat = 46.897122, zoom = 4 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
 addPolylines(data = roads, color = "purple", weight = 1, opacity = 0.7) %>% 
  addCircles(lng = milestone_EDCS_coord$Longitude, 
             lat = milestone_EDCS_coord$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "blue" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", milestone_EDCS_coord$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", milestone_EDCS_coord$place,
                           "<br><b> Type of inscription: </b>", milestone_EDCS_coord$inscr_type,
                              "<br><b> Material: </b>", milestone_EDCS_coord$Material,
                              "<br><b> Not before (date): </b>", milestone_EDCS_coord$start_yr,
                             "<br><b> Not after (date): </b>", milestone_EDCS_coord$end_yr_1,
                                "<br><b> Commentary: </b>", milestone_EDCS_coord$Comment),
             ) %>% 
addLegend(position = "topright",
  colors = c("Blue", "Purple"),
  labels = c("Milestones", "Roman roads (BA atlas)"), opacity = 1,
  title = "Milestones and Roman roads" 
)
map_milestones_EDCS_roads
```

## Location of milestones compared with all inscriptions

```{r}
# inscriptions with coordinates

EDCS_tibble$Latitude <- as.numeric(EDCS_tibble$Latitude)
EDCS_tibble$Longitude <- as.numeric(EDCS_tibble$Longitude)

```

```{r}
# how many inscriptions of the entire dataset have no coordinates
EDCS_no_coords<- EDCS_tibble %>% 
  filter(is.na(EDCS_tibble$Latitude & is.na(EDCS_tibble$Longitude)))
EDCS_no_coords

# how many inscriptions of the entire dataset do have coordinates
EDCS_has_coords<- EDCS_tibble %>% 
  filter(!is.na(EDCS_tibble$Latitude | !is.na(EDCS_tibble$Longitude)))
EDCS_has_coords

# how many % of inscriptions with coordinates is represented by milestones 
nrow(milestone_EDCS_coord)/(nrow(EDCS_has_coords)/100)
```

```{r}
map_milestones_EDCS_compared<- leaflet(width="100%") %>%
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
# addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
# addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
# addPolylines(data = roads, color = "purple", weight = 2, opacity = 0.7) %>% 
  setView( lng = 22.326743, lat = 46.897122, zoom = 4 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>% 
  addCircles(lng = EDCS_tibble$Longitude, 
             lat = EDCS_tibble$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "grey" , fillColor = "white",) %>% 
  addCircles(lng = milestone_EDCS_coord$Longitude, 
             lat = milestone_EDCS_coord$Latitude, opacity = 0.5, radius = 5, fill = TRUE, color = "purple" , fillColor = "",
             popup = paste0("<b> InscriptionID: </b>", milestone_EDCS_coord$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", milestone_EDCS_coord$place,
                           "<br><b> Type of inscription: </b>", milestone_EDCS_coord$inscr_type,
                              "<br><b> Material: </b>", milestone_EDCS_coord$Material,
                              "<br><b> Not before (date): </b>", milestone_EDCS_coord$start_yr,
                             "<br><b> Not after (date): </b>", milestone_EDCS_coord$end_yr_1,
                                "<br><b> Commentary: </b>", milestone_EDCS_coord$Comment),
             ) %>% 
addLegend(position = "topright",
  colors = c("Purple", "Grey"),
  labels = c("Milestones (1.51 %)", "Other inscriptions (98.49 %)"), opacity = 1,
  title = "Milestones vs other inscriptions (EDCS) with known coordinates"
)
map_milestones_EDCS_compared
```


# Temporal exploration
```{r}
dates<- milestone_EDCS %>% 
  dplyr::select(start_yr, end_yr_1)

dates$start_yr <- as.numeric(dates$start_yr)
dates$end_yr_1 <- as.numeric(dates$end_yr_1)

# basic descriptive overview 
summary(dates$start_yr)
summary(dates$end_yr_1)
```


## Temporal distribution of 'start_yr' years of milestones
```{r}
plot(table(dates$start_yr), ylab= "Number of inscriptions", xlab="'Terminus post quem' in years")
```

## Temporal distribution of 'end_yr_1' years of milestones
```{r}
plot(table(dates$end_yr_1), ylab= "Number of inscriptions", xlab="'Terminus ante quem' in years")
```


## Interval
How precisely were the milestones dated? Shorter the date interval is, the more precisely was the inscription dated.

```{r}
dates<- dates %>%  
  mutate(interval = abs(end_yr_1-start_yr)) 
```

```{r}
dates %>% 
  ggplot(aes(x=interval)) +
  geom_histogram(fill="black", size=10, binwidth = 10) +
  labs(x = "Length of date interval", y = "Number of inscription", title = "The length of date interval (start_yr - end_yr_1)", subtitle = "n = 3737" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(0,25, 50, 75, 100,200, 300, 400, 500, 600))
```

### What is the basic statistics for the date interval?
```{r}
summary(dates$interval)
```
The milestones are one of the most precisely dated types of inscriptions, with majority dated with precision to 1 year.

## Mean value of the date interval
```{r}
dates<- dates %>%  
  mutate(mean = rowMeans(dates[,1:2]))
```

How many milestones contains non-finite temporal data (containin one or two NAs)
```{r}
sum(is.na(dates$mean))
```

### Plotting the result in 10 year bins
```{r}
dates %>% 
  ggplot(aes(x=mean)) +
  geom_histogram(fill="orange", size=1, binwidth = 10) +
  labs(x = "Date (ar. mean)", y = "Number of inscription", title = "Number of dated milestones in time (ar. mean) in ten year bins", subtitle = "n = 3737" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(-100,1,100,200, 250, 300, 400))+
  geom_vline(aes(xintercept=250))
```
```{r}
ggsave("../output/milestones_mean_10yrs.png", width = 10, height= 6)
```

### Plotting the result in 50 year bins
```{r}
dates %>% 
  ggplot(aes(x=mean)) +
  geom_histogram(fill="orange", size=1, binwidth = 50) +
  labs(x = "Date (ar. mean)", y = "Number of inscription", title = "Number of dated milestones in time (ar. mean) in 50 year bins", subtitle = "n = 3737" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(-100,1,100,200, 250, 300, 400))+
  geom_vline(aes(xintercept=250))
```


### Plotting the result in 100 year bins
```{r}
dates %>% 
  ggplot(aes(x=mean)) +
  geom_histogram(fill="orange", size=1, binwidth = 100) +
  labs(x = "Date (ar. mean)", y = "Number of inscription", title = "Number of dated milestones in time (ar. mean) in 100 year bins", subtitle = "n = 3737" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(-100,1,100,200, 250, 300, 400))+
  geom_vline(aes(xintercept=250))
```


---

*The explorations will continue by:*

- spatiotemporal visualisations > spatial distributions of milestones over time
- machine learning & typological definitions > digitally enhaced exploration of what makes milestone a milestone
- generating a set of guidelines for future epigraphers and students on how to recognize a milestone
- searching for new milestones in digital datasets, applying the newly created set of rules of what defines a milestone to the rest of EDCS and other large datasets of inscriptions

**Any questions, suggestions, and comments are more than welcome at `petra.hermankova@cas.au.dk`.**





