---
title: "EDH Milestones"
author: "Petra Hermankova"
date: "01/07/2020"
output: html_document:
  df_print: paged
    
---

```{r setup, echo=TRUE, message=FALSE}
devtools::install_github("mplex/cedhar", subdir="pkg/sdam")
library(tidyverse)
library(tidytext)
library(dplyr)
library(stringr)
library(sdam)
library(jsonlite)
library(getPass)
library(formatR)
```

### Loading data

1. Input your sciencedata.dk username - type directly into the RStudio console
```{r, echo = FALSE }
user <- readline("your sciencedata username: ")
```

2. Make the request (you will be asked for password in a new pop-up window)
```{r, echo = FALSE }
resp = request("EDH_cleaned_2020-06-26.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDH/", method="GET", cred=c(user, getPass("your sciencedata password: ")))
```

3. Make a list and tibble from the request function
```{r}
list_json <- jsonlite::fromJSON(resp)
EDH_tibble = as_tibble(list_json)
```

# Milestones as type of an inscription

```{r}
unique(EDH_tibble$type_of_inscription_clean)
```
```{r}
milestone_insc <- EDH_tibble %>% 
  filter(type_of_inscription_clean == "mile-/leaguestone")
```

```{r}
milestone_insc %>% 
  count(milestone_insc$type_of_monument_clean, sort = TRUE)
```

# Milestones as type of monument

```{r}
milestone_monument <- EDH_tibble %>% 
  filter(EDH_tibble$type_of_monument_clean == "mile-/leaguestone")
```

```{r}
milestone_monument %>% 
  count(milestone_monument$type_of_inscription_clean, sort = TRUE)
```

```{r}
EDH_tibble %>% 
  names()
```


```{r}
milestone_combined<- EDH_tibble %>% 
  filter(type_of_monument_clean == "mile-/leaguestone" | type_of_inscription_clean == "mile-/leaguestone") 
```

## Distribution by provinces

```{r}
milestone_combined %>% 
  count(province_label_clean, sort =TRUE)
```

```{r, fig.height=10, fig.width=10}
milestone_combined %>% 
  count(province_label_clean, sort =TRUE) %>% 
  ggplot(aes(y=province_label_clean, x=n), size = n) +
  geom_point(color="red") +
  coord_fixed(ratio = 7/1)
  
```

## Dates
```{r}
dates<- milestone_combined %>% 
  select(not_before, not_after, origdate_text, origdate_attrs) 
dates_flat <- flatten(dates)
dates_flat$not_before <- as.numeric(dates_flat$not_before)
dates_flat$not_after <- as.numeric(dates_flat$not_after)
summary(dates_flat$not_before)
```

```{r}
plot(dates_flat$not_before)
```

```{r}
summary(dates_flat$not_after)
```
```{r}
plot(dates_flat$not_after)
```

# Temporal distribution of 'not_before' years of milestones
```{r}
plot(table(dates_flat$not_before), ylab= "Number of inscriptions", xlab="'Not before' YEAR")
```

## Combining dates and provinces - unfinished

```{r,chunk-label, results='hide', fig.height=4}
milestones_flat <- flatten(milestone_combined)
milestones_flat
```


