---
title: "Exploration of milestone inscriptions in the EDH dataset"
author: "Petra Hermankova"
date: "01/10/2020"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged 
 
---

This notebook explores all the milestones in the **Epigraphic Database Heidelberg** (https://edh-www.adw.uni-heidelberg.de/, henceforth EDH). The primary aim is to:

- find all inscriptions categorized as 'milestones'
- provide basic descriptive statistics: how many there are, what categories of monuments we have, what materials they are made from, where do they come from etc. 
- display all the milestones on the map along with the map of Roman provinces and Roman roads

The script takes the reader through all the steps of selection, descriptive statistics and through the various displays on the interactive map. Address all the questions and comments to `petra.hermankova@cas.au.dk`.

# Initial setup

## Setup of the environment:

```{r setup, echo=TRUE, message=FALSE}
devtools::install_github("sdam-au/sdam") # loading SDAM custom package, if not working try devtools::install_github("mplex/cedhar", subdir="pkg/sdam")
#devtools::install_github("mplex/cedhar", subdir="pkg/sdam")
library(tidyverse)
library(sdam)
library(jsonlite)
library(getPass)
library(formatR)
library(leaflet)
```

## Loading data
Load the dataset, if you have Sciencedata.dk credentials

```{r, echo=FALSE}
mycred_secret<- readLines("~/mysecret.txt")
```

```{r, loading data}
resp = request("EDH_attrs_cleaned_2020-09-30.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDH/public", method="GET", cred=mycred_secret)
```

```{r, echo=FALSE}
remove(mycred_secret)
```

OR of you don't have Sciencedata credentials:
```{r, loading data public}
resp = request("EDH_attrs_cleaned_2020-09-30.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDH/public", method="GET")
```

Make a list and tibble from the downloaded dataset
```{r}
list_json <- jsonlite::fromJSON(resp)
EDH_tibble = as_tibble(list_json)
```

## Exploring the data
Display the first six records of the dataset
```{r}
head(EDH_tibble)
```

# Finding all the milestones 
What attributes (columns) might contain milestone related information? Let's display all the the attribute names.
```{r}
EDH_tibble %>% 
  names()
```
Based on the experience and previous explorations of the dataset, any milestone related information is most likely contained in the following attributes: `type_of_inscription_clean`, `type_of_monumnet_clean`, `commentary`,`support_objecttype`, `support_objecttype_text`, `keywords_term` and `keywords_term_text`.


## Milestone as type of inscription

```{r}
unique(EDH_tibble$type_of_inscription_clean)
```
### How many inscriptions are categorised as a milestone?
```{r}
milestone_insc <- EDH_tibble %>% 
  filter(type_of_inscription_clean == "mile-/leaguestone")
nrow(milestone_insc)
```

## Milestone as type of monument

```{r}
unique(EDH_tibble$type_of_monument_clean)
```
### How many inscriptions are categorised as milestone?
```{r}
milestone_monument <- EDH_tibble %>% 
  filter(EDH_tibble$type_of_monument_clean == "mile-/leaguestone")

nrow(milestone_monument)
```

## Milestone in the `commentary` attribute
Inspect the commentary for milestone related comments

### How many inscriptions contain any milestone-related commentary?
```{r}
milestone_comment<- EDH_tibble %>%  
  dplyr::filter(str_detect(commentary, "[M|m]eilen"))
nrow(milestone_comment)
```
### Explore the first six commentaries mentioning milestones
```{r}
head(milestone_comment$commentary)
```

## Milestone in the `support_objecttype` attribute 
The attribute `support_object_type` represents code for the Eagle Europeana standard vocabulary (https://www.eagle-network.eu/voc/objtyp/lod/89.html), where the Milestone has a LOD code 89.

How many there are?
```{r}
milestone_lod_object<- EDH_tibble %>% 
  filter(support_objecttype =="89") 
nrow(milestone_lod_object)
```
## Milestone in the `support_objecttype_text` attribute
Checking the consistency of `support_objecttype` and `support_objecttype_text` attributes to see if the attributes refer to the same type.

```{r}
EDH_tibble %>% 
  select(support_objecttype, support_objecttype_text) %>% 
  dplyr::filter(str_detect(support_objecttype_text, "[M|m]eilen")) %>% 
  count(support_objecttype_text)
```
The results are consistent and the free text corresponds with the LOD.


## Milestone in the `keywords_term` attribute
The attribute `keywords_term` represents code for the Eagle Europeana standard vocabulary (https://www.eagle-network.eu/voc/typeins/lod/102.html), where the Milestone has a LOD code 102.

How many there are?
```{r}
milestone_lod_insctype<- EDH_tibble %>% 
  filter(keywords_term =="102") 
nrow(milestone_lod_insctype)
```

## Milestone in the `support_objecttype_text` attribute
Checking the consistency of `keywords_term` and `keywords_term_text` attributes to see if the attributes refer to the same type.

```{r}
EDH_tibble %>% 
  select(keywords_term, keywords_term_text) %>% 
  dplyr::filter(str_detect(keywords_term_text, "[M|m]eilen")) %>% 
  count(keywords_term_text)
```
The results are consistent and the free text corresponds with the LOD.


## Combining all milestones

Now we will combine all milestones that have a word meilen- in `commentary` and are defined as milestones in the following attributes: `type_of_inscription_clean`, `type_of_monumnet_clean`,`support_objecttype`, `keywords_term`.

```{r}
milestone_all <- EDH_tibble %>%  
  dplyr::filter(str_detect(commentary, "[M|m]eilen") | type_of_monument_clean == "mile-/leaguestone" | type_of_inscription_clean == "mile-/leaguestone" | support_objecttype =="89" | keywords_term =="102") 
```


# Descriptive overview 

## How many milestones total are in the EDH dataset when we combine all the search criteria?
```{r}
nrow(milestone_all)
```

## How many % of the total number of EDH inscriptions represent milestones
```{r}
nrow(milestone_all)/(nrow(EDH_tibble)/100)
```

## Type of inscription

What different types of inscriptions belong to the discussed group of milestones?
```{r}
milestone_all %>% 
  count(milestone_all$type_of_inscription_clean, sort = TRUE)
```
## Type of monument

What different types of monuments belong to the discussed group of milestones?
```{r}
milestone_all %>% 
  count(milestone_all$type_of_monument_clean, sort = TRUE)
```

## Language

What language were written the inscriptions belonging to the discussed group of milestones?
```{r}
unnest_auto(milestone_all, language) %>% 
  count(language, sort=TRUE)
```


## Material 
What different types of material is represented in the discussed group of milestones?
```{r}
milestone_all %>% 
  count(material_clean, sort=TRUE)
```

## Dimensions
What are the dimensions of the discussed group of milestones?
```{r}
milestone_dims <- milestone_all %>% 
  dplyr::select(height_cm, width_cm, depth_cm)
milestone_dims <- as.data.frame(milestone_dims)
```

### Basic statistical summary of Height in cm
An overview statistics for all Heights.
```{r}
summary(milestone_dims$height_cm)
```

### Basic statistical summary of Width in cm
An overview statistics for all Widths.
```{r}
summary(milestone_dims$width_cm)
```

### Basic statistical summary of Depth in cm
An overview statistics for all Depths.
```{r}
summary(milestone_dims$depth_cm)
```

# Spatial exploration

## Ancient findspot
What kind of information do we have about ancient findspot of a milestone? Let's see first 20 records.
```{r}
milestone_all$findspot_ancient_clean[1:20]
```

How many different ancient findspots do we have, including "NULL"?
```{r}
length(unique(milestone_all$findspot_ancient_clean))
```

## Distribution by Roman provinces
How many milestones were found in individual Roman provinces? Display from the province containing the most to least milestones.
```{r}
milestone_all %>% 
  count(province_label_clean, sort =TRUE)
```

Display all the provinces and milestones in a chart: 
```{r, fig.height=10, fig.width=7}
milestone_all %>% 
  count(province_label_clean, sort =TRUE) %>% 
  ggplot(aes(y=province_label_clean, x=n), size = n) +
  geom_point(color="red") +
  coord_fixed(ratio = 7/1) +
  labs(x = "Number of milestones", y = "Roman province", title = "Number of milestones per Roman Province", subtitle = "n = 1770" ) +
    theme_linedraw(base_size = 12)
```

## Map of all milestones (with available coordinates)

```{r}
# selecting milestones with coordinates 
coords_milestone<- as.data.frame(cbind(id = milestone_all$id, 
                                       coordinates = milestone_all$coordinates, 
                                       findspot_ancient_clean = milestone_all$findspot_ancient_clean,
                                       type_of_inscription_clean = milestone_all$type_of_inscription_clean, 
                                       type_of_monument_clean = milestone_all$type_of_monument_clean, 
                                       not_before = milestone_all$not_before,
                                       not_after = milestone_all$not_after,
                                       commentary = milestone_all$commentary))

# milestones with no coordinates
coords_milestone_empty<- coords_milestone %>% 
  dplyr::filter(coordinates == "list()")

# milestones with coordinates
coords_milestone_full<- coords_milestone %>% 
  dplyr::filter(coordinates != "list()")

lat_long_milestone<- coords_milestone_full %>% 
  separate(col = coordinates, into = c("longitude", "latitude"), sep = ",")

lat_long_milestone$latitude <- as.numeric(str_replace(lat_long_milestone$latitude, pattern = "\\)", replacement=""))
lat_long_milestone$longitude <- as.numeric(str_replace(lat_long_milestone$longitude, pattern = "c\\(", replacement=""))
```

### How many milestones have spatial coordinates
```{r}
nrow(lat_long_milestone)
```

## Milestones as dots
```{r}
map_milestones_dot <-leaflet(width="100%") %>%
  addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
  addCircles(lng = lat_long_milestone$longitude, 
             lat = lat_long_milestone$latitude, radius = 10, fill = TRUE, color= , fillColor = lat_long_milestone$type_of_monument_clean,
             popup = paste0("<b> InscriptionID: </b>", lat_long_milestone$id, 
                            "<br><b> Ancient findspot: </b>", lat_long_milestone$findspot_ancient,
                             "<br><b> Type of inscription: </b>", lat_long_milestone$type_of_inscription_clean,
                              "<br><b> Type of monument: </b>", lat_long_milestone$type_of_monument_clean,
                              "<br><b> Not before (date): </b>", lat_long_milestone$not_before,
                              "<br><b> Not after (date): </b>", lat_long_milestone$not_after,
                                "<br><b> Commentary: </b>", lat_long_milestone$commentary),
) %>% 
addLegend(position = "topright",
  colors = c("Blue"),
  labels = c("Milestones (n=1770)"), opacity = 1,
  title = "Position of all milestones (EDH)" 
)
map_milestones_dot

```


## Clustering of milestones
```{r}
map_milestones_cluster <-leaflet(width="100%") %>%
  #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
  addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  addMarkers(lng = lat_long_milestone$longitude, 
             lat = lat_long_milestone$latitude,
             popup = paste0("<b> InscriptionID: </b>", lat_long_milestone$id, 
                            "<br><b> Ancient findspot: </b>", lat_long_milestone$findspot_ancient,
                             "<br><b> Type of inscription: </b>", lat_long_milestone$type_of_inscription_clean,
                              "<br><b> Type of monument: </b>", lat_long_milestone$type_of_monument_clean,
                              "<br><b> Not before (date): </b>", lat_long_milestone$not_before,
                                "<br><b> Not after (date): </b>", lat_long_milestone$not_after,
                                "<br><b> Commentary: </b>", lat_long_milestone$commentary),
             
    clusterOptions = markerClusterOptions()
  ) %>% 
addLegend(position = "topright",
  colors = c("Blue", "Green", "Yellow", "Orange"),
  labels = c("Individual inscription", "Small cluster", "Medium cluster", "Large cluster"), opacity = 1,
  title = "Clustering of all milestones in the EDH dataset" 
)
map_milestones_cluster
```

## Milestones and Roman provinces (200 AD)

```{r, echo=FALSE}
library(rgdal)
library(raster)
library(sf)
library(sp)
```
source Roads: http://awmc.unc.edu/awmc/map_data/shapefiles/ba_roads/
source Roman provinces (AD 117): http://awmc.unc.edu/awmc/map_data/shapefiles/cultural_data/political_shading/roman_empire_ad_117/shape/

```{r, echo=FALSE}
setwd("~/Github/epigraphic_roads")
dir.create("../data/ba_roads")
unzip("../data/ba_roads.zip", exdir = "../data/ba_roads/")
dir.create("../data/roman_empire_200_ad_provinces")
unzip("../data/roman_empire_ad_200_provinces.zip", exdir = "../data/roman_empire_200_ad_provinces")
```

```{r, echo=FALSE}
roads <- st_read("../data/ba_roads/ba_roads.shp")
provinces200 <- st_read("../data/roman_empire_200_ad_provinces/roman_empire_ad_200_provinces.shp")
```

```{r, echo=FALSE}
orbis <- readLines("https://raw.githubusercontent.com/emeeks/orbis_v2/master/base_routes.geojson") %>% paste(collapse = "\n") %>% 
  fromJSON(simplifyVector = FALSE)
```

```{r}

map_milestones_provinces<- leaflet(width="100%") %>%
 #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
 #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
 addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
 addPolylines(data = provinces200, color = "red", weight = 2, opacity = 0.7) %>% 
  addCircles(lng = lat_long_milestone$longitude, 
             lat = lat_long_milestone$latitude, radius = 10, fill = TRUE, color = "blue" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", lat_long_milestone$id, 
                            "<br><b> Ancient findspot: </b>", lat_long_milestone$findspot_ancient,
                             "<br><b> Type of inscription: </b>", lat_long_milestone$type_of_inscription_clean,
                              "<br><b> Type of monument: </b>", lat_long_milestone$type_of_monument_clean,
                              "<br><b> Not before (date): </b>", lat_long_milestone$not_before,
                              "<br><b> Not after (date): </b>", lat_long_milestone$not_after,
                                "<br><b> Commentary: </b>", lat_long_milestone$commentary),
) %>% 
addLegend(position = "topright",
  colors = c("Blue", "Red"),
  labels = c("Milestones", "Roman Provinces (200 AD)"), opacity = 1,
  title = "Milestones in the Roman Empire" 
)
map_milestones_provinces
```

## Milestones and Roman roads
```{r}
map_milestones_roads<- leaflet(width="100%") %>%
 addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
 #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
 #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  setView( lng = 22.326743, lat = 46.897122, zoom = 4 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
 addPolylines(data = roads, color = "purple", weight = 1, opacity = 0.7) %>% 
  addCircles(lng = lat_long_milestone$longitude, 
             lat = lat_long_milestone$latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "blue" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", lat_long_milestone$id, 
                            "<br><b> Ancient findspot: </b>", lat_long_milestone$findspot_ancient,
                             "<br><b> Type of inscription: </b>", lat_long_milestone$type_of_inscription_clean,
                              "<br><b> Type of monument: </b>", lat_long_milestone$type_of_monument_clean,
                              "<br><b> Not before (date): </b>", lat_long_milestone$not_before,
                              "<br><b> Not after (date): </b>", lat_long_milestone$not_after,
                                "<br><b> Commentary: </b>", lat_long_milestone$commentary),
             ) %>% 
addLegend(position = "topright",
  colors = c("Blue", "Purple"),
  labels = c("Milestones", "Roman roads (BA atlas)"), opacity = 1,
  title = "Milestones and Roman roads" 
)
map_milestones_roads
```

## Location of milestones compared with all inscriptions

```{r}
# inscriptions with coordinates

lat_long_EDH <- EDH_tibble %>%  
  separate(col = coordinates, into = c("longitude", "latitude"), sep = ",")
lat_long_EDH

lat_long_EDH$latitude <- as.numeric(str_replace(lat_long_EDH$latitude, pattern = "\\)", replacement=""))
lat_long_EDH$longitude <- as.numeric(str_replace(lat_long_EDH$longitude, pattern = "c\\(", replacement=""))

```
```{r}
# how many inscriptions of the entire dataset have coordinates
EDH_no_coords<- sum(is.na(lat_long_EDH$latitude))
EDH_no_coords

# how many inscriptions of the entire dataset do not have coordinates
EDH_has_coords<- sum(!is.na(lat_long_EDH$latitude))
EDH_has_coords
# how many % of the entire dataset milestones represent 
nrow(lat_long_milestone)/(nrow(lat_long_EDH)/100)

# how many % of inscriptions with coordinates is represented by milestones 
nrow(lat_long_milestone)/(EDH_has_coords/100)
```

```{r}
map_milestones_compared<- leaflet(width="100%") %>%
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
# addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
# addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
# addPolylines(data = roads, color = "purple", weight = 2, opacity = 0.7) %>% 
  setView( lng = 22.326743, lat = 46.897122, zoom = 4 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>% 
  addCircles(lng = lat_long_EDH$longitude, 
             lat = lat_long_EDH$latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "grey" , fillColor = "white",) %>% 
  addCircles(lng = lat_long_milestone$longitude, 
             lat = lat_long_milestone$latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "purple" , fillColor = "",
             popup = paste0("<b> InscriptionID: </b>", lat_long_milestone$id, 
                            "<br><b> Ancient findspot: </b>", lat_long_milestone$findspot_ancient,
                             "<br><b> Type of inscription: </b>", lat_long_milestone$type_of_inscription_clean,
                              "<br><b> Type of monument: </b>", lat_long_milestone$type_of_monument_clean,
                              "<br><b> Not before (date): </b>", lat_long_milestone$not_before,
                              "<br><b> Not after (date): </b>", lat_long_milestone$not_after,
                                "<br><b> Commentary: </b>", lat_long_milestone$commentary),
             ) %>% 
addLegend(position = "topright",
  colors = c("Purple", "Grey"),
  labels = c("Milestones (2.24 %)", "Other inscriptions (97.76 %)"), opacity = 1,
  title = "Milestones vs other inscriptions (EDH)"
)
map_milestones_compared
```


# Temporal exploration
```{r}
dates<- milestone_all %>% 
  dplyr::select(not_before, not_after)

dates$not_before <- as.numeric(dates$not_before)
dates$not_after <- as.numeric(dates$not_after)

# basic descriptive overview 
summary(dates$not_before)
summary(dates$not_after)
```


## Temporal distribution of 'not_before' years of milestones
```{r}
plot(table(dates$not_before), ylab= "Number of inscriptions", xlab="'Not before' YEAR")
```

## Temporal distribution of 'not_after' years of milestones
```{r}
plot(table(dates$not_after), ylab= "Number of inscriptions", xlab="'Not after' YEAR")
```
## Interval
How precisely were the milestones dated? Shorter the date interval is, the more precisely was the inscription dated.

```{r}
dates<- dates %>%  
  mutate(interval = (not_before - not_after) *-1)
```
```{r}
dates %>% 
  ggplot(aes(x=interval)) +
  geom_histogram(fill="black", size=10, binwidth = 10) +
  labs(x = "Length of date interval", y = "Number of inscription", title = "The length of date interval (not_before - not_after)", subtitle = "n = 1172" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(1,100,200, 300, 400, 500, 600))
```

### What is the basic statistics for the date interval?
```{r}
summary(dates$interval)
```
The milestones are one of the most precisely dated types of inscriptions, with majority dated with precision 3 years or less.

## Mean value of the date interval
```{r}
dates<- dates %>%  
  mutate(mean = rowMeans(dates[,1:2]))
```
How many milestones contains non-finite temporal data (containin one or two NAs)
```{r}
sum(is.na(dates$mean))
```

### Plotting the result in 10 year bins
```{r}
dates %>% 
  ggplot(aes(x=mean)) +
  geom_histogram(fill="orange", size=1, binwidth = 10) +
  labs(x = "Date (ar. mean)", y = "Number of inscription", title = "Number of dated milestones in time (ar. mean) in ten year bins", subtitle = "n = 1172" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(-100,1,100,200, 250, 300, 400))+
  geom_vline(aes(xintercept=250))
```
```{r}
ggsave("milestones_mean_10yrs.png", width = 10, height= 6)
```

### Plotting the result in 50 year bins
```{r}
dates %>% 
  ggplot(aes(x=mean)) +
  geom_histogram(fill="orange", size=1, binwidth = 50) +
  labs(x = "Date (ar. mean)", y = "Number of inscription", title = "Number of dated milestones in time (ar. mean) in 50 year bins", subtitle = "n = 1172" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(-100,1,100,200, 250, 300, 400))+
  geom_vline(aes(xintercept=250))
```


### Plotting the result in 100 year bins
```{r}
dates %>% 
  ggplot(aes(x=mean)) +
  geom_histogram(fill="orange", size=1, binwidth = 100) +
  labs(x = "Date (ar. mean)", y = "Number of inscription", title = "Number of dated milestones in time (ar. mean) in 50 year bins", subtitle = "n = 1172" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(-100,1,100,200, 250, 300, 400))+
  geom_vline(aes(xintercept=250))
```


---

*The explorations will continue by:*

- spatiotemporal visualisations > spatial distributions of milestones over time
- machine learning & typological definitions > digitally enhaced exploration of what makes milestone a milestone
- generating a set of guidelines for future epigraphers and students on how to recognize a milestone
- searching for new milestones in digital datasets, applying the newly created set of rules of what defines a milestone to the rest of EDH and other large datasets of inscriptions

**Any questions, suggestions, and comments are more than welcome at `petra.hermankova@cas.au.dk`.**



