---
title: "Road related inscriptions from Bithynia and Sardinia"
author: "Petra Hermankova"
date: "25/03/2021"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged 
---

# Initial setup

## Setup of the environment:

```{r setup, echo=TRUE, message=FALSE}
devtools::install_github("sdam-au/sdam") # loading SDAM custom package, if not working try devtools::install_github("mplex/cEDCSar", subdir="pkg/sdam")
#devtools::install_github("mplex/cEDCSar", subdir="pkg/sdam")
library(tidyverse)
library(sdam)
library(jsonlite)
library(getPass)
library(formatR)
library(leaflet)
```

This notebook explores all the milestones from Roman provinces **Bithynia** and **Sardinia** in the **Epigraphic Database Clauss-Slaby** (http://www.manfredclauss.de/, henceforth EDCS). The primary aim is to:

- find all inscriptions categorized as 'milestones'
- provide basic descriptive statistics: how many there are, what categories of monuments we have, what materials they are made from, where do they come from etc. 
- display all the milestones on the map along with the map of Roman provinces and Roman roads

The script takes the reader through all the steps of selection, descriptive statistics and through the various displays on the interactive map. Address all the questions and comments to `petra.hermankova@cas.au.dk`.

## Loading data
Load the dataset, if you have Sciencedata.dk credentials

```{r, echo=FALSE}
mycred_secret<- readLines("~/mysecret.txt")
```

```{r, loading data}
resp2 = request("EDCS_text_cleaned_2021-03-01.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDCS/public", method="GET", cred=mycred_secret)
```

OR anonymously:
```{r, loading data public, eval=FALSE}
resp2 = request("EDCS_text_cleaned_2021-03-01.json", path="/public/1f5f56d09903fe259c0906add8b3a55e/", method="GET", anonymous = TRUE, cred = NULL)
```

Make a list and tibble from the downloaded dataset
```{r}
list_json2 <- jsonlite::fromJSON(resp2)
EDCS_tibble = as_tibble(list_json2)
```

```{r}
remove(list_json2, resp2)
```

```{r}
names(EDCS_tibble)
```

# Finding all milestones in Bithynia and Sardinia
```{r}
Bithynia_milestones<- EDCS_tibble %>% 
  filter(inscr_type %in% str_subset(inscr_type, "miliaria")) %>% 
  filter(province == "Pontus et Bithynia" ) %>% 
  select(`EDCS-ID`, province, place, start_yr, end_yr_1, inscr_type, inscription, clean_text_interpretive_word, language, Latitude, Longitude, Material) %>% 
  unnest_wider(inscr_type) %>% 
  rename("inscr_type1" = "...1", "inscr_type2" = "...2","inscr_type3" = "...3")


Sardinia_milestones<- EDCS_tibble %>% 
  filter(inscr_type %in% str_subset(inscr_type, "miliaria")) %>% 
  filter(province == "Sardinia") %>% 
  select(`EDCS-ID`, province, place, start_yr, end_yr_1, inscr_type, inscription, clean_text_interpretive_word, language, Latitude, Longitude, Material) %>% 
  unnest_wider(inscr_type) %>% 
  rename("inscr_type1" = "...1", "inscr_type2" = "...2","inscr_type3" = "...3")

```

## Export as CSV
```{r}
getwd()
write.csv(Bithynia_milestones, "../output/Milestones_EDCS_Bithynia.csv", row.names = T)
write.csv(Sardinia_milestones, "../output/Milestones_EDCS_Sardinia.csv", row.names = T)
```

### All funerary inscriptions in Bithynia and Sardinia

```{r}
funerary_Bithynia<- EDCS_tibble %>% 
  filter(province == "Pontus et Bithynia") %>%
  filter(inscr_type %in% str_subset(inscr_type, "tituli sepulcrales")) %>% 
  select(`EDCS-ID`, province, place, start_yr, end_yr_1, inscr_type, inscription, clean_text_interpretive_word, language, Latitude, Longitude, Material) %>% 
  unnest_wider(inscr_type) %>% 
  rename("inscr_type1" = "...1", "inscr_type2" = "...2","inscr_type3" = "...3")

funerary_Bithynia
```

```{r}
funerary_Sardinia<- EDCS_tibble %>% 
  filter(province == "Sardinia") %>%
  filter(inscr_type %in% str_subset(inscr_type, "tituli sepulcrales")) %>% 
  select(`EDCS-ID`, province, place, start_yr, end_yr_1, inscr_type, inscription, clean_text_interpretive_word, language, Latitude, Longitude, Material) %>% 
  unnest_wider(inscr_type) %>% 
  rename("inscr_type1" = "...1", "inscr_type2" = "...2","inscr_type3" = "...3","inscr_type4" = "...4","inscr_type5" = "...5")

funerary_Sardinia
```

## Export as CSV
```{r}
getwd()
write.csv(funerary_Sardinia, "../output/funerary_EDCS_Sardinia.csv", row.names = T)
write.csv(funerary_Bithynia, "../output/funerary_EDCS_Bithynia.csv", row.names = T)
```


# Inscriptions with road-related terms

```{r, loading data public terms, eval=FALSE}
#resp3 = request("EDCS_terms_2021-03-25.json", path="/public/66cbabddae0e02c6ae6c15be9746990c/", method="GET", anonymous = TRUE, cred = NULL)
resp3 = request("EDCS_terms_2021-03-25.json", path="/public/66cbabddae0e02c6ae6c15be9746990c/", method="GET", cred=mycred_secret)

list_json3 <- jsonlite::fromJSON(resp3, flatten = TRUE, simplifyDataFrame=T)
EDCS_vocabs <- as_tibble(list_json3)
```

```{r}
remove(list_json3, resp3, mycred_secret)
```

```{r}
head(EDCS_vocabs)
EDCS_vocabs %>% 
  filter(terms_N != 0) -> EDCS_road_terms
```

```{r}
road_json<- toJSON(EDCS_road_terms, auto_unbox = TRUE)
road_list<- fromJSON(road_json)
EDCS_road = as_tibble(road_list)
EDCS_road
```

Cleaning up memory:
```{r}
remove(EDCS_road_terms, road_json, road_list)
```


## Road related isncriptions in Sardinia
```{r}
road_Sardinia<- EDCS_road %>% 
  filter(province == "Sardinia") %>% 
  select(`EDCS-ID`, province, place, start_yr, end_yr_1, inscr_type, inscription, clean_text_interpretive_word, language, Latitude, Longitude, Material, terms, terms_N) %>% 
  unnest_wider(inscr_type) %>% 
  rename("inscr_type1" = "...1", "inscr_type2" = "...2","inscr_type3" = "...3","inscr_type4" = "...4") %>% 
  unnest_wider(terms) %>% 
  rename("term1" = "...1", "term2" = "...2") %>% 
  unnest_wider(start_yr) %>% rename("start_yr" = "...1") %>% 
  unnest_wider(end_yr_1) %>% rename("end_yr_1" = "...1") %>% 
  unnest_wider(language) %>% rename("language" = "...1") %>% 
  unnest_wider(Latitude) %>% rename("Latitude" = "...1") %>% 
  unnest_wider(Longitude) %>% rename("Longitude" = "...1") %>% 
  unnest_wider(Material) %>% rename("Material" = "...1")

```
## Road related isncriptions in Bithynia
```{r}
road_Bithynia<- EDCS_road %>% 
  filter(province == "Pontus et Bithynia") %>% 
  select(`EDCS-ID`, province, place, start_yr, end_yr_1, inscr_type, inscription, clean_text_interpretive_word, language, Latitude, Longitude, Material, terms, terms_N) %>% 
  unnest_wider(inscr_type) %>% 
  rename("inscr_type1" = "...1", "inscr_type2" = "...2","inscr_type3" = "...3") %>% 
  unnest_wider(terms) %>% 
  rename("term1" = "...1", "term2" = "...2") %>% 
  unnest_wider(start_yr) %>% rename("start_yr" = "...1") %>% 
  unnest_wider(end_yr_1) %>% rename("end_yr_1" = "...1") %>% 
  unnest_wider(language) %>% rename("language" = "...1") %>% 
  unnest_wider(Latitude) %>% rename("Latitude" = "...1") %>% 
  unnest_wider(Longitude) %>% rename("Longitude" = "...1") %>% 
  unnest_wider(Material) %>% rename("Material" = "...1")
```

## Export as CSV
```{r}
getwd()
write.csv(road_Sardinia, "../output/Road_terms_EDCS_Sardinia.csv", row.names = T)
write.csv(road_Bithynia, "../output/Road_terms_EDCS_Bithynia.csv", row.names = T)
```




# Maps with roads

```{r, echo=FALSE}
library(rgdal)
library(raster)
library(sf)
library(sp)
```
Source of Roads shapefile: http://awmc.unc.edu/awmc/map_data/shapefiles/ba_roads/
Source of Roman provinces (AD 200): http://awmc.unc.edu/awmc/map_data/shapefiles/cultural_data/political_shading/roman_empire_ad_200/shape/

```{r, echo=FALSE}
setwd("~/Github/epigraphic_roads")
dir.create("../data/ba_roads")
unzip("../data/ba_roads.zip", exdir = "../data/ba_roads/")
dir.create("../data/roman_empire_200_ad_provinces")
unzip("../data/roman_empire_ad_200_provinces.zip", exdir = "../data/roman_empire_200_ad_provinces")
```

```{r, echo=FALSE}
roads <- st_read("../data/ba_roads/ba_roads.shp")
provinces200 <- st_read("../data/roman_empire_200_ad_provinces/roman_empire_ad_200_provinces.shp")
```

```{r, echo=FALSE}
orbis <- readLines("https://raw.githubusercontent.com/emeeks/orbis_v2/master/base_routes.geojson") %>% paste(collapse = "\n") %>% 
  fromJSON(simplifyVector = FALSE)
```

```{r}
Bithynia_milestones$Latitude <- as.numeric(Bithynia_milestones$Latitude)
Bithynia_milestones$Longitude <- as.numeric(Bithynia_milestones$Longitude)

funerary_Bithynia$Latitude <- as.numeric(funerary_Bithynia$Latitude)
funerary_Bithynia$Longitude <- as.numeric(funerary_Bithynia$Longitude)

road_Bithynia$Latitude <- as.numeric(road_Bithynia$Latitude)
road_Bithynia$Longitude <- as.numeric(road_Bithynia$Longitude)
```



# Bithynia
```{r}
map_Bithynia_EDCS<- leaflet(width="100%") %>%
 #addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
 #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
 #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  setView( lng = 31.6110479, lat = 40.7332953, , zoom = 6 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
 addPolylines(data = roads, color = "purple", weight = 1, opacity = 0.7) %>% 
  addCircles(lng = Bithynia_milestones$Longitude, 
             lat = Bithynia_milestones$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "blue" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", Bithynia_milestones$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", Bithynia_milestones$place,
                           "<br><b> Type of inscription: </b>", Bithynia_milestones$inscr_type,
                              "<br><b> Material: </b>", Bithynia_milestones$Material,
                              "<br><b> Not before (date): </b>", Bithynia_milestones$start_yr,
                             "<br><b> Not after (date): </b>", Bithynia_milestones$end_yr_1,
                                "<br><b> Commentary: </b>", Bithynia_milestones$Comment),
             ) %>% 
  addCircles(lng = funerary_Bithynia$Longitude, 
             lat = funerary_Bithynia$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "orange" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", funerary_Bithynia$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", funerary_Bithynia$place,
                           "<br><b> Type of inscription: </b>", funerary_Bithynia$inscr_type,
                              "<br><b> Material: </b>", funerary_Bithynia$Material,
                              "<br><b> Not before (date): </b>", funerary_Bithynia$start_yr,
                             "<br><b> Not after (date): </b>", funerary_Bithynia$end_yr_1,
                                "<br><b> Commentary: </b>", funerary_Bithynia$Comment),
             ) %>% 
  addCircles(lng = road_Bithynia$Longitude, 
             lat = road_Bithynia$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "yellow" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", road_Bithynia$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", road_Bithynia$place,
                           "<br><b> Type of inscription: </b>", road_Bithynia$inscr_type,
                              "<br><b> Material: </b>", road_Bithynia$Material,
                              "<br><b> Not before (date): </b>", road_Bithynia$start_yr,
                             "<br><b> Not after (date): </b>", road_Bithynia$end_yr_1,
                                "<br><b> Commentary: </b>", road_Bithynia$Comment),
             ) %>%
addLegend(position = "topright",
  colors = c("Blue", "Orange", "Yellow", "Purple"),
  labels = c("Milestones", "Funerary insc", "Road related insc", "Roman roads (BA atlas)"), opacity = 1,
  title = "Road related inscriptions in Bythinia" 
)
map_Bithynia_EDCS

```

# Sardinia
```{r}
Sardinia_milestones$Latitude <- as.numeric(Sardinia_milestones$Latitude)
Sardinia_milestones$Longitude <- as.numeric(Sardinia_milestones$Longitude)

funerary_Sardinia$Latitude <- as.numeric(funerary_Sardinia$Latitude)
funerary_Sardinia$Longitude <- as.numeric(funerary_Sardinia$Longitude)

road_Sardinia$Latitude <- as.numeric(road_Sardinia$Latitude)
road_Sardinia$Longitude <- as.numeric(road_Sardinia$Longitude)
```


```{r}
map_Sardinia_EDCS<- leaflet(width="100%") %>%
 #addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
 #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
 #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  setView( lng = 9.087933, lat = 40.070653, , zoom = 7 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
 addPolylines(data = roads, color = "purple", weight = 1, opacity = 0.7) %>% 
  addCircles(lng = Sardinia_milestones$Longitude, 
             lat = Sardinia_milestones$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "blue" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", Sardinia_milestones$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", Sardinia_milestones$place,
                           "<br><b> Type of inscription: </b>", Sardinia_milestones$inscr_type,
                              "<br><b> Material: </b>", Sardinia_milestones$Material,
                              "<br><b> Not before (date): </b>", Sardinia_milestones$start_yr,
                             "<br><b> Not after (date): </b>", Sardinia_milestones$end_yr_1,
                                "<br><b> Commentary: </b>", Sardinia_milestones$Comment),
             ) %>% 
  addCircles(lng = funerary_Sardinia$Longitude, 
             lat = funerary_Sardinia$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "orange" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", funerary_Sardinia$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", funerary_Sardinia$place,
                           "<br><b> Type of inscription: </b>", funerary_Sardinia$inscr_type,
                              "<br><b> Material: </b>", funerary_Sardinia$Material,
                              "<br><b> Not before (date): </b>", funerary_Sardinia$start_yr,
                             "<br><b> Not after (date): </b>", funerary_Sardinia$end_yr_1,
                                "<br><b> Commentary: </b>", funerary_Sardinia$Comment),
             ) %>% 
  addCircles(lng = road_Sardinia$Longitude, 
             lat = road_Sardinia$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "yellow" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", road_Sardinia$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", road_Sardinia$place,
                           "<br><b> Type of inscription: </b>", road_Sardinia$inscr_type,
                              "<br><b> Material: </b>", road_Sardinia$Material,
                              "<br><b> Not before (date): </b>", road_Sardinia$start_yr,
                             "<br><b> Not after (date): </b>", road_Sardinia$end_yr_1,
                                "<br><b> Commentary: </b>", road_Sardinia$Comment),
             ) %>%
addLegend(position = "topright",
  colors = c("Blue", "Orange", "Yellow", "Purple"),
  labels = c("Milestones", "Funerary insc", "Road related insc", "Roman roads (BA atlas)"), opacity = 1,
  title = "Road related inscriptions in Bythinia" 
)
map_Sardinia_EDCS

```
```{r}
library(oceanis)
export_png(map_Sardinia_EDCS, "../output/", "Inscriptions_EDCS_Sardinia")
export_png(map_Bithynia_EDCS, "../output/", "Inscriptions_EDCS_Bithynia")
```

