---
title: "Milestones around Decapolis"
author: "Petra Hermankova"
date: "26/03/2021"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged 
---

# Initial setup

## Setup of the environment:

```{r setup, echo=TRUE, message=FALSE}
devtools::install_github("sdam-au/sdam") # loading SDAM custom package, if not working try devtools::install_github("mplex/cEDCSar", subdir="pkg/sdam")
#devtools::install_github("mplex/cEDCSar", subdir="pkg/sdam")
library(tidyverse)
library(sdam)
library(jsonlite)
library(getPass)
library(formatR)
library(leaflet)
```

This notebook explores all the milestones from the Roman Decapolis (cities Gerasa, Gadara, Hippos, Dion, Abila, Capitolias, Nysa-Scythopolis, Pella ,Philadelphia) in Roman provinces **Arabia**, **Syria** and **Palaestina** in the **Epigraphic Database Clauss-Slaby** (http://www.manfredclauss.de/, henceforth EDCS). The primary aim is to:

- find all inscriptions categorized as 'milestones'
- provide basic descriptive statistics: how many there are, what categories of monuments we have, what materials they are made from, where do they come from etc. 
- display all the milestones on the map along with the map of Roman provinces and Roman roads

The script takes the reader through all the steps of selection, descriptive statistics and through the various displays on the interactive map. Address all the questions and comments to `petra.hermankova@cas.au.dk`.

## Loading data
Load the dataset, if you have Sciencedata.dk credentials

```{r, echo=FALSE}
mycred_secret<- readLines("~/mysecret.txt")
```

```{r, loading data}
resp2 = request("EDCS_text_cleaned_2021-03-01.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDCS/public", method="GET", cred=mycred_secret)
```

OR anonymously:
```{r, loading data public, eval=FALSE}
resp2 = request("EDCS_text_cleaned_2021-03-01.json", path="/public/1f5f56d09903fe259c0906add8b3a55e/", method="GET", anonymous = TRUE, cred = NULL)
```

Make a list and tibble from the downloaded dataset
```{r}
list_json2 <- jsonlite::fromJSON(resp2)
EDCS_tibble = as_tibble(list_json2)
```

# Decapolis

```{r}
decapolis <- c("Gerasa", "Gadara", "Hippos", "Dion", "Abila", "Capitolias", "Nysa-Scythopolis", "Pella" , "Philadelphia")

decapolis_province <- c("Syria", "Palaestina", "Arabia")



decapolis_inscriptions<- EDCS_tibble %>% 
  filter(province %in% str_subset(province, decapolis_province)) %>% 
  filter(place %in% str_subset(place, decapolis)) 

decapolis_inscriptions %>% 
  count(province) %>% 
  View()


provinces_inscriptions<- EDCS_tibble %>% 
  filter(province %in% str_subset(province, decapolis_province)) 

```

```{r}
milestones<- provinces_inscriptions %>% 
  filter(inscr_type %in% str_subset(inscr_type, "miliaria"))

termini<- provinces_inscriptions %>% 
  filter(inscr_type %in% str_subset(inscr_type, "termini"))
```
## How many inscriptions, milestones and termini?
```{r}
nrow(provinces_inscriptions)
nrow(milestones)
nrow(termini)
```
## In which province?
```{r}
provinces_inscriptions %>% 
  count(province, sort=T)
```

```{r}
milestones %>% 
  count(province, sort=T)
```

```{r}
termini %>% 
  count(province, sort=T)
```

## When

```{r}
dates_all<- provinces_inscriptions %>% 
  mutate(mean = rowMeans(provinces_inscriptions[,22:23]))
```
### Dated all inscriptions 

```{r}
dates_all %>% 
  ggplot(aes(x=mean)) +
  geom_histogram(fill="orange", size=1, binwidth = 10) +
  labs(x = "Date (ar. mean)", y = "Number of inscription", title = "Number of dated inscriptions in time (ar. mean) in ten year bins", subtitle = "n = 3351" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(-100,1,100,200, 250, 300, 400))+
  geom_vline(aes(xintercept=106))
```

#### Inscriptions dated before 106 BC
```{r}
dates_all %>% 
  filter(mean < 106) 
```


### Dated Milestones 
```{r}
dates_milestone<- milestones %>%  
  mutate(mean = rowMeans(milestones[,22:23]))

dates_milestone %>% 
  ggplot(aes(x=mean)) +
  geom_histogram(fill="orange", size=1, binwidth = 10) +
  labs(x = "Date (ar. mean)", y = "Number of inscription", title = "Number of dated milestones in time (ar. mean) in ten year bins", subtitle = "n = 1038" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(-100,1,100,200, 250, 300, 400))+
  geom_vline(aes(xintercept=106))
```

#### Filter milestones before 106 AD
```{r}
dates_milestone %>% 
  filter(mean <106)
```


### Dated Termini

```{r}
dates_termini<- termini %>%  
  mutate(mean = rowMeans(termini[,22:23]))

dates_termini %>% 
  ggplot(aes(x=mean)) +
  geom_histogram(fill="orange", size=1, binwidth = 10) +
  labs(x = "Date (ar. mean)", y = "Number of inscription", title = "Number of dated termini in time (ar. mean) in ten year bins", subtitle = "n = 227" ) +
  theme_linedraw(base_size = 12) +
  scale_x_discrete(limits=c(0,50, 100,150, 200, 250))+
  geom_vline(aes(xintercept=106))
```

#### Filter termini before 106 AD
```{r}
dates_termini %>% 
  filter(mean ==61)
```




# Maps

```{r, echo=FALSE}
library(rgdal)
library(raster)
library(sf)
library(sp)
```
Source of Roads shapefile: http://awmc.unc.edu/awmc/map_data/shapefiles/ba_roads/
Source of Roman provinces (AD 200): http://awmc.unc.edu/awmc/map_data/shapefiles/cultural_data/political_shading/roman_empire_ad_200/shape/

```{r, echo=FALSE}
setwd("~/Github/epigraphic_roads")
dir.create("../data/ba_roads")
unzip("../data/ba_roads.zip", exdir = "../data/ba_roads/")
dir.create("../data/roman_empire_200_ad_provinces")
unzip("../data/roman_empire_ad_200_provinces.zip", exdir = "../data/roman_empire_200_ad_provinces")
```

```{r, echo=FALSE}
roads <- st_read("../data/ba_roads/ba_roads.shp")
provinces200 <- st_read("../data/roman_empire_200_ad_provinces/roman_empire_ad_200_provinces.shp")
```

```{r, echo=FALSE}
orbis <- readLines("https://raw.githubusercontent.com/emeeks/orbis_v2/master/base_routes.geojson") %>% paste(collapse = "\n") %>% 
  fromJSON(simplifyVector = FALSE)
```



# Map all inscriptions from provinces Syria, Palaestina, Arabia

```{r}
library(raster)
library(sf)

provinces_inscriptions$Latitude <- as.numeric(provinces_inscriptions$Latitude)
provinces_inscriptions$Longitude <- as.numeric(provinces_inscriptions$Longitude)

decapolis_coords<- as.data.frame(read.csv("../data/decapolis/Decapolis.csv"))
decapolis_coords

```

```{r}

library(leaflet)
map_provinces_inscriptions_EDCS<- leaflet(width="100%") %>%
 #addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
 #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
 #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  setView( lng = 35.9239625, lat = 31.9515694, zoom = 5 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
  addPolylines(data = roads, color = "purple", weight = 1, opacity = 0.7) %>% 
 addCircles(lng = provinces_inscriptions$Longitude, 
             lat = provinces_inscriptions$Latitude, opacity = 0.2, radius = 10, fill = TRUE, color = "blue" , fillColor = "black",
             popup = paste0("<b> InscriptionID: </b>", provinces_inscriptions$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", provinces_inscriptions$place,
                           "<br><b> Type of inscription: </b>", provinces_inscriptions$inscr_type,
                              "<br><b> Material: </b>", provinces_inscriptions$Material,
                              "<br><b> Not before (date): </b>", provinces_inscriptions$start_yr,
                            "<br><b> Not after (date): </b>", provinces_inscriptions$end_yr_1,
                                "<br><b> Commentary: </b>", provinces_inscriptions$Comment),
             ) %>% 
  addCircles(lng = decapolis_coords$X, 
             lat = decapolis_coords$Y,
             opacity = 0.5, radius = 15, fill = TRUE, color = "red" , fillColor = "black", popup = paste0("<b> City: </b>", decapolis_coords$Name)) %>% 

addLegend(position = "bottomright",
  colors = c("Blue", "Purple", "Red"),
  labels = c("Inscriptions", "Roman roads (BA atlas)", "Decapolis cities"), opacity = 1,
  title = "Inscriptions in provinces Syria, Palaestina, Arabia" 
) %>% 
  addScaleBar(position="bottomleft")
map_provinces_inscriptions_EDCS

```


## Map of milestones
```{r}
milestones$Latitude <- as.numeric(milestones$Latitude)
milestones$Longitude <- as.numeric(milestones$Longitude)
```

```{r}
library(leaflet)
map_milestones_EDCS<- leaflet(width="100%") %>%
 #addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
 #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
 #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  setView( lng = 35.9239625, lat = 31.9515694, zoom = 5 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
 addPolylines(data = roads, color = "purple", weight = 1, opacity = 0.7) %>% 
  addCircles(lng = milestones$Longitude, 
             lat = milestones$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "blue" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", milestones$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", milestones$place,
                           "<br><b> Type of inscription: </b>", milestones$inscr_type,
                              "<br><b> Material: </b>", milestones$Material,
                              "<br><b> Not before (date): </b>", milestones$start_yr,
                             "<br><b> Not after (date): </b>", milestones$end_yr_1,
                                "<br><b> Commentary: </b>", milestones$Comment),
             ) %>% 
  addCircles(lng = decapolis_coords$X, 
             lat = decapolis_coords$Y,
             opacity = 0.5, radius = 15, fill = TRUE, color = "red" , fillColor = "black", popup = paste0("<b> City: </b>", decapolis_coords$Name)) %>% 
addLegend(position = "bottomright",
  colors = c("Blue", "Purple", "Red"),
  labels = c("Inscriptions", "Roman roads (BA atlas)", "Decapolis cities"), opacity = 1,
  title = "Milestones in provinces Syria, Palaestina, Arabia" 
) %>% 
  addScaleBar(position="bottomleft")

map_milestones_EDCS

```

# Milestones with keywords 

## Caput viae
```{r}
milestones %>% 
  filter(clean_text_interpretive_word %in% str_subset(milestones$clean_text_interpretive_word, "caput")) 
```

```{r}
milestones %>% 
  filter(clean_text_interpretive_word %in% str_subset(milestones$clean_text_interpretive_word, "^viae")) 
```

## Ab [city]
```{r}
milestones %>% 
  filter(clean_text_interpretive_word %in% str_subset(milestones$clean_text_interpretive_word, "ab ")) 
```

## Milia passuum and variants
```{r}
milestones %>% 
  filter(clean_text_interpretive_word %in% str_subset(milestones$clean_text_interpretive_word, "passu*")) 
```

```{r}
milestones %>% 
  filter(clean_text_interpretive_word %in% str_subset(milestones$clean_text_interpretive_word, "^mill*")) 
```

```{r}
milestones %>% 
  filter(clean_text_interpretive_word %in% str_subset(milestones$clean_text_interpretive_word, "milia")) 
```

# Text 

## Milestone

```{r}
unique(milestones$clean_text_interpretive_word)
```


## Termini

```{r}
unique(termini$clean_text_interpretive_word)
```



# Export as CSV

```{r}
inscriptions_EDCS<- provinces_inscriptions %>% 
  dplyr::select(`EDCS-ID`, province, place, start_yr, end_yr_1, inscr_type, inscription, clean_text_interpretive_word, language, Latitude, Longitude, Material) %>% 
  unnest_wider(inscr_type) %>% 
  rename("inscr_type1" = "...1", "inscr_type2" = "...2","inscr_type3" = "...3")

write.csv(inscriptions_EDCS, "../output/Inscriptions_all_decapolis_EDCS.csv", row.names = T)
```

