---
title: "Inscriptions from Bithynia (EDCS)"
author: "Petra Hermankova"
date: "03/05/2021"
output:
  word_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 2
    toc_float: yes
---

# Initial setup

## Setup of the environment:

```{r setup, echo=TRUE, message=FALSE}
devtools::install_github("sdam-au/sdam") # loading SDAM custom package, if not working try devtools::install_github("mplex/cEDCSar", subdir="pkg/sdam")
#devtools::install_github("mplex/cEDCSar", subdir="pkg/sdam")
library(tidyverse)
library(sdam)
library(jsonlite)
library(getPass)
library(formatR)
library(leaflet)
```

This notebook explores all the milestones from Roman provinces **Bithynia** in the **Epigraphic Database Clauss-Slaby** (http://www.manfredclauss.de/, henceforth EDCS). The primary aim is to:

- find all inscriptions categorized as 'milestones'
- provide basic descriptive statistics: how many there are, what categories of monuments we have, what materials they are made from, where do they come from etc. 
- display all the milestones on the map along with the map of Roman provinces and Roman roads

The script takes the reader through all the steps of selection, descriptive statistics and through the various displays on the interactive map. Address all the questions and comments to `petra.hermankova@cas.au.dk`.

## Loading data

Load the dataset, if you have Sciencedata.dk credentials

```{r, echo=FALSE}
mycred_secret<- readLines("~/mysecret.txt")
```

```{r, loading data}
resp2 = request("EDCS_text_cleaned_2021-03-01.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDCS/public", method="GET", cred=mycred_secret)
```

```{r, loading data public, eval=FALSE}
resp2 = request("EDCS_text_cleaned_2021-03-01.json", path="/public/1f5f56d09903fe259c0906add8b3a55e/", method="GET", anonymous = TRUE, cred = NULL)
```

Make a list and tibble from the downloaded dataset
```{r}
EDCS_tibble <- as_tibble(jsonlite::fromJSON(resp2))
```

# All inscriptions in EDCS

```{r}
head(EDCS_tibble, 5)
```

```{r}
names(EDCS_tibble)
```

# Inscriptions with road-related terms

```{r, loading data public terms, eval=FALSE}
resp3 = request("EDCS_terms_2021-03-25.json", path="/public/66cbabddae0e02c6ae6c15be9746990c/", method="GET", anonymous = TRUE, cred = NULL)
#resp3 = request("EDCS_terms_2021-03-25.json", path="/public/66cbabddae0e02c6ae6c15be9746990c/", method="GET", cred=mycred_secret)

EDCS_vocabs <- as_tibble(jsonlite::fromJSON(resp3))
```

```{r}
EDCS_road_terms<- EDCS_vocabs %>% 
  filter(terms_N != 0)
```

```{r}
road_json<- toJSON(EDCS_road_terms, auto_unbox = TRUE)
EDCS_road<- as_tibble(fromJSON(road_json))
EDCS_road
```


## Finding all milestones in Bithynia 
```{r}
Bithynia_milestones<- EDCS_tibble %>% 
  filter(inscr_type %in% str_subset(inscr_type, "miliaria")) %>% 
  filter(province == "Pontus et Bithynia" ) %>% 
  select(`EDCS-ID`, province, place, start_yr, end_yr_1, inscr_type, inscription, clean_text_interpretive_word, language, Latitude, Longitude, Material) %>% 
  unnest_wider(inscr_type) %>% 
  rename("inscr_type1" = "...1", "inscr_type2" = "...2","inscr_type3" = "...3")

Bithynia_milestones
```


## Funerary inscriptions in Bithynia 

```{r}
funerary_Bithynia<- EDCS_tibble %>% 
  filter(province == "Pontus et Bithynia") %>%
  filter(inscr_type %in% str_subset(inscr_type, "tituli sepulcrales")) %>% 
  select(`EDCS-ID`, province, place, start_yr, end_yr_1, inscr_type, inscription, clean_text_interpretive_word, language, Latitude, Longitude, Material) %>% 
  unnest_wider(inscr_type) %>% 
  rename("inscr_type1" = "...1", "inscr_type2" = "...2","inscr_type3" = "...3")

funerary_Bithynia
```

## Road related inscriptions in Bithynia
```{r}
road_Bithynia<- EDCS_road %>% 
  filter(province == "Pontus et Bithynia") %>% 
  select(`EDCS-ID`, province, place, start_yr, end_yr_1, inscr_type, inscription, clean_text_interpretive_word, language, Latitude, Longitude, Material, terms, terms_N) %>% 
  unnest_wider(inscr_type) %>% 
  rename("inscr_type1" = "...1", "inscr_type2" = "...2","inscr_type3" = "...3") %>% 
  unnest_wider(terms) %>% 
  rename("term1" = "...1", "term2" = "...2") %>% 
  unnest_wider(start_yr) %>% rename("start_yr" = "...1") %>% 
  unnest_wider(end_yr_1) %>% rename("end_yr_1" = "...1") %>% 
  unnest_wider(language) %>% rename("language" = "...1") %>% 
  unnest_wider(Latitude) %>% rename("Latitude" = "...1") %>% 
  unnest_wider(Longitude) %>% rename("Longitude" = "...1") %>% 
  unnest_wider(Material) %>% rename("Material" = "...1")

road_Bithynia
```



# Maps with roads

```{r, echo=FALSE}
library(rgdal)
library(raster)
library(sf)
library(sp)
```
Source of Roads shapefile: http://awmc.unc.edu/awmc/map_data/shapefiles/ba_roads/
Source of Roman provinces (AD 200): http://awmc.unc.edu/awmc/map_data/shapefiles/cultural_data/political_shading/roman_empire_ad_200/shape/

```{r, echo=FALSE}
setwd("~/Github/epigraphic_roads")
dir.create("../data/ba_roads")
unzip("../data/ba_roads.zip", exdir = "../data/ba_roads/")
dir.create("../data/roman_empire_200_ad_provinces")
unzip("../data/roman_empire_ad_200_provinces.zip", exdir = "../data/roman_empire_200_ad_provinces")
```

```{r, echo=FALSE}
roads <- st_read("../data/ba_roads/ba_roads.shp")
provinces200 <- st_read("../data/roman_empire_200_ad_provinces/roman_empire_ad_200_provinces.shp")
```

```{r, echo=FALSE}
orbis <- readLines("https://raw.githubusercontent.com/emeeks/orbis_v2/master/base_routes.geojson") %>% paste(collapse = "\n") %>% 
  fromJSON(simplifyVector = FALSE)
```

```{r}
Bithynia_milestones$Latitude <- as.numeric(Bithynia_milestones$Latitude)
Bithynia_milestones$Longitude <- as.numeric(Bithynia_milestones$Longitude)

funerary_Bithynia$Latitude <- as.numeric(funerary_Bithynia$Latitude)
funerary_Bithynia$Longitude <- as.numeric(funerary_Bithynia$Longitude)

road_Bithynia$Latitude <- as.numeric(road_Bithynia$Latitude)
road_Bithynia$Longitude <- as.numeric(road_Bithynia$Longitude)
```



# Bithynia
```{r}
map_Bithynia_EDCS<- leaflet(width="100%") %>%
 #addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
 #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
 #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  setView( lng =  31.6075, lat = 40.734722 , zoom = 7 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
 addPolylines(data = roads, color = "purple", weight = 1, opacity = 0.7) %>% 
  addCircles(lng = Bithynia_milestones$Longitude, 
             lat = Bithynia_milestones$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "blue" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", Bithynia_milestones$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", Bithynia_milestones$place,
                           "<br><b> Type of inscription: </b>", Bithynia_milestones$inscr_type,
                              "<br><b> Material: </b>", Bithynia_milestones$Material,
                              "<br><b> Not before (date): </b>", Bithynia_milestones$start_yr,
                             "<br><b> Not after (date): </b>", Bithynia_milestones$end_yr_1,
                                "<br><b> Commentary: </b>", Bithynia_milestones$Comment),
             ) %>% 
  addCircles(lng = funerary_Bithynia$Longitude, 
             lat = funerary_Bithynia$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "orange" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", funerary_Bithynia$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", funerary_Bithynia$place,
                           "<br><b> Type of inscription: </b>", funerary_Bithynia$inscr_type,
                              "<br><b> Material: </b>", funerary_Bithynia$Material,
                              "<br><b> Not before (date): </b>", funerary_Bithynia$start_yr,
                             "<br><b> Not after (date): </b>", funerary_Bithynia$end_yr_1,
                                "<br><b> Commentary: </b>", funerary_Bithynia$Comment),
             ) %>% 
  addCircles(lng = road_Bithynia$Longitude, 
             lat = road_Bithynia$Latitude, opacity = 0.5, radius = 10, fill = TRUE, color = "yellow" , fillColor = "red",
             popup = paste0("<b> InscriptionID: </b>", road_Bithynia$`EDCS-ID`, 
                            "<br><b> Ancient findspot: </b>", road_Bithynia$place,
                           "<br><b> Type of inscription: </b>", road_Bithynia$inscr_type,
                              "<br><b> Material: </b>", road_Bithynia$Material,
                              "<br><b> Not before (date): </b>", road_Bithynia$start_yr,
                             "<br><b> Not after (date): </b>", road_Bithynia$end_yr_1,
                                "<br><b> Commentary: </b>", road_Bithynia$Comment),
             ) %>%
addLegend(position = "bottomleft",
  colors = c("Blue", "Orange", "Yellow", "Purple"),
  labels = c("Milestones", "Funerary insc", "Road related insc", "Roman roads (BA atlas)"), opacity = 1,
  title = "Road related inscriptions in Bithynia (EDCS)" 
)
map_Bithynia_EDCS

```